---
title: "Cointegration Analysis of Token Pairs"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This report evaluates the cointegration relationships between token pairs from the provided dataset. The results are visualized as a heatmap, showing confidence levels for cointegration. Note that confidence levels rely on correctly identifying unit roots in the original series. Without this verification, the cointegration test may yield misleading results.

```{r load-libraries}
# Load necessary libraries
library(urca)        # For cointegration test
library(tidyverse)   # For data manipulation and visualization
library(reshape2)    # For reshaping data
```

```{r load-data}
# Load the dataset
file_path <- "/Users/michael/Desktop/pairs-trading-bot/price data/5minutedatatest.csv"  # Update with the correct file path if necessary
data <- read.csv(file_path)

# Ensure Date column is correctly formatted
data$Date <- as.Date(data$Date)

# Remove rows with missing values
cleaned_data <- data %>% drop_na()

# Preview the data
head(cleaned_data)
```

```{r prepare-data}
# Get token names
tokens <- colnames(cleaned_data)[-1]  # Exclude the Date column

# Initialize a results dataframe
results <- data.frame(
  Token1 = character(),
  Token2 = character(),
  p_value = numeric(),
  confidence_level = character(),
  stringsAsFactors = FALSE
)
```

```{r test-cointegration}
# Perform Engle-Granger cointegration test for each pair
for (i in 1:(length(tokens) - 1)) {
  for (j in (i + 1):length(tokens)) {
    token1 <- tokens[i]
    token2 <- tokens[j]

    # Extract time series for the pair
    ts1 <- cleaned_data[[token1]]
    ts2 <- cleaned_data[[token2]]

    # Perform linear regression
    lm_model <- lm(ts1 ~ ts2)
    residuals <- lm_model$residuals

    # Perform the Augmented Dickey-Fuller test on residuals
    adf_test <- ur.df(residuals, type = "none", selectlags = "AIC")
    p_value <- summary(adf_test)@testreg$coefficients[1, 4]

    # Determine confidence level
    if (p_value < 0.01) {
      confidence <- "99%"
    } else if (p_value < 0.05) {
      confidence <- "95%"
    } else if (p_value < 0.10) {
      confidence <- "90%"
    } else {
      confidence <- "None"
    }

    # Append to results
    results <- rbind(results, data.frame(Token1 = token1, Token2 = token2, p_value = p_value, confidence_level = confidence))
  }
}

# Preview the results
dim(results)
head(results)
```

```{r create-heatmap}
# Prepare data for heatmap
heatmap_data <- results %>%
  filter(confidence_level != "None") %>%
  select(Token1, Token2, confidence_level) %>%
  mutate(value = case_when(
    confidence_level == "99%" ~ 3,
    confidence_level == "95%" ~ 2,
    confidence_level == "90%" ~ 1,
    TRUE ~ 0
  ))

# Reshape data for heatmap
heatmap_matrix <- dcast(heatmap_data, Token1 ~ Token2, value.var = "value")
heatmap_matrix[is.na(heatmap_matrix)] <- 0

# Create heatmap
heatmap_long <- melt(heatmap_matrix, id.vars = "Token1")
colnames(heatmap_long) <- c("Token1", "Token2", "Confidence")

# Plot the heatmap
ggplot(heatmap_long, aes(x = Token1, y = Token2, fill = factor(Confidence))) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c("#F0F0F0", "#ADD8E6", "#6495ED", "#00008B"),
                    labels = c("None", "90%", "95%", "99%"),
                    name = "Confidence Level") +
  labs(title = "Cointegration Heatmap",
       x = "Token 1",
       y = "Token 2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r extract-results}
# Extract unique pairs with 99% confidence level
cointegrated_pairs <- results %>%
  filter(confidence_level == "99%") %>%
  select(Token1, Token2) %>%
  distinct()

# Print the cointegrated pairs
print(cointegrated_pairs)

```

